\newpage
\section*{Abstract}
The main objective of this project is to develop an intelligent parking system that combines sensor technology and cloud and edge computing power. A comprehensive study of the state of the art in this area has been carried out, investigating the different types of sensors available and their applicability in parking environments.

Field tests were carried out to evaluate the accuracy and reliability of the sensors in detecting parking space availability. These tests allowed data to be collected on space occupancy in real-time and to evaluate sensor performance under real conditions.

In addition, a mobile application was designed and developed using Flutter, a cross-platform development framework, which allows users to access real-time information on parking space availability. The application offers features such as searching for nearby parking and reserving a space in it.

To ensure optimal performance and fast response, computing power was leveraged both in the cloud and edge computing for processing sensor data. Using a Raspberry Pi, data sent by sensors via LoRa could be efficiently processed and real-time updates provided to application users.

In summary, this project covered everything from state-of-the-art analysis to the construction of an intelligent parking system (prototype). Suitable sensors were investigated, field tests were conducted, a mobile application was developed, and cloud and edge computing solutions were implemented to offer a complete user experience. Finally, a section of possible improvements was developed.
